A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2016-06-29 Answer_to_referee_report_LensingPaper.tex

\item The manuscript in its current form is also slightly longer than necessary at certain
points: for instance, in section 3.2.1, it is not really necessary to present all
the lensing equations (e.g. the time delay, equation 7), and similarly, section
3.3.1 contains a lot of material that is implicit in Cappellari's JAM code, and
therefore does not need to be explained equation by equation here. Section 3.3.4
also rather throws the book at the NFW profile relations – it probably isn't
necessary to include equation 22. \Comment{We made an effort to remove unnecessary equations in Section 3.2.1, tighten the explanation of the derivation of the Jeans equations in Section 3.3.1 and 3.3.3 and removed big and clumsy mathematical expressions by referring to the explicit equations in the literature instead. In 3.3.2 we refer stronger to the changes we made to Capellari's JAM code and kept the corresponding equations. In Section 3.3.4 we moved equations that are definitions and therefore needed inline to make it shorter; we kept the Maccio concentration vs. halo mass relation and explicitely mention that we will use this later as a prior in our modelling. In general we include a lot of equations and a big theory introduction because it is a paper that is potentially interesting for two communities, the lensing and the dynamics community. We wanted to make the paper understandable for both.}

\item From a scientific perspective, in section 4.3.3, the authors comment that the
varying M/L JAM model is unable to reproduce the dynamics when M/L is forced to be
always rising as a function of radius, but it would be interesting to see if the
model could be improved by removing this constraint. Perhaps the authors could add
such a model to this section, or else comment on why they do not attempt it. \Comment{We have indeed tried a ``mass-follows-light'' JAM model with constant anisotropy and free M/L profile early on in our investigations of J1331. This was unsuccessful to a degree that showing the result in the paper would not be reasonable.  The way how we can generate variable M/L profiles with our MGE mass model is restricted. We have 6 data points and 5 fit M/L parameter within 5 arcsec. This lead to an over-fitting in the inner regions where most of the Gaussians are, while not having enough flexibility at larger radii. (We added a sentence mentioning that in the paper.) The resulting profile was very wiggly, difficult to interpret and most likely unphysical.  As we lay out in the discussion, constraints on a varying stellar M/L ratio should be motivated by stellar population analysis in the future.}

### 2016-06-30 Answer_to_referee_report_LensingPaper.tex

\item From a scientific perspective, in section 4.3.3, the authors comment that the
varying M/L JAM model is unable to reproduce the dynamics when M/L is forced to be
always rising as a function of radius, but it would be interesting to see if the
model could be improved by removing this constraint. Perhaps the authors could add
such a model to this section, or else comment on why they do not attempt it. \Comment{We have indeed tried a ``mass-follows-light'' JAM model with constant anisotropy and free M/L profile early on in our investigations of J1331. This was unsuccessful to a degree that showing the result in the paper would not be reasonable.  The way how we can generate variable M/L profiles with our MGE mass model is restricted. We have 6 data points and 5 fit M/L parameter within 5 arcsec. This lead to an over-fitting in the inner regions where most of the Gaussians are, while not having enough flexibility at larger radii. (We added a sentence mentioning that in the paper.) The resulting profile was very wiggly, difficult to interpret and most likely unphysical.  As we lay out in the discussion, constraints on a varying stellar M/L ratio should be motivated by stellar population analysis in the future.}

\item In section 4.4.1, the NFW profile is fitted only to the inner data, and then is
shown to not provide a good model for the data beyond that point. This is not
entirely surprising, given that the outer data were not used to constrain the model!
What happens if the NFW profile is fitted to both small- and large- radius data at
once? (Perhaps it is impossible to find a good NFW model across the whole radius
range, but if this is the case, then it should be stated.) \Comment{Yes, you are completely right. Our initial tests tried of course to fit the whole range of data, but failed. The problem is that we were never able to fit both the central $v_\text{rms}$ dip and the dip around $R'\sim 6''$ with any of the models we tried. If you think this is frustrating, we definitely feel the same way. Our best explanation was, that the spiral arms do weird things to the kinematics around $R'\sim 5-7''$ that cannot be captured by the axisymmtric Jeans models. In this region our MGE models also were not a good fit to the galaxy anymore. Considering also that there are other studies focussing on the outer regions and the counter-rotating core is the most interesting feature, we decided to focus our efforts on the inner regions and to try to explain the central kinematics only. We added however a small new section 4.4.4 that discusses that and why modelling attempts to the whole data range fail and what possible next modelling steps should be.}

\item The discussion section is too non-committal and speculative in parts. For instance,
in section 5.1, the M/Ls inferred from a number of the different models are compared
with M/Ls from the literature, but the authors do not make any value judgement
between the different models to come to any firmer conclusions about what they
believe the most convincing description of the IMF to be. At the moment, this
section seems to just be trying to show that their various models can be made to
agree with previous models from the literature, but it would be much more
interesting if some attempt could be made to resolve all the different models (both
in this paper and in previous works) or at least present them in a more coherent
framework, to try to come up with a better picture of the nature of J1331's IMF. \Comment{This is a valid point of criticism. We thought again about the different models and indications for different IMFs that we collected and came to the conclusion that an IMF slightly less bottom-beavy than the Salpeter IMF is the most likely model for J1331's bulge. We tried to give Section 5.1 (now 5.2) more structure and tweaked the discussion to present arguments for $\Upsilon_{I,*} \sim 4.2$ and and to argue against the alternative models.\\We also moved the section that discusses J1331's merger history to the beginning to the discussion section, focussed more on the what the merger could have done to J1331's mass distribution and less on modelling failures. All following sections use the background information about mergers to discuss our models.\\Our scientific honesty is however, that without further data, we cannot be more commital and some of our statements have to stay in the "speculation regime".}

\item In Section 5.2, the authors could make suggestions for what might be causing the two
$v_{rms}$ dips. Currently, they just state what cannot be the cause, but do not make any
more convincing suggestions – e.g. could it be associated with the possible recent
merger that is explored in Section 5.3? It would be more satisfying if some
suggestions like this could be put forward. \Comment{Good point. We admit that the text was formulated a bit too negative, and that we can do better in explaining J1331 based on our models than just ruling out things. Also, some of the effects that mergers can have on galaxies \emph{could}, if they occured in J1331, actually explain some of the peculiar kinematic features of J1331. We tried to re-formulate section 5.2 (now 5.3) and discuss several aspects of the kinematics with respect to the merger history.}

### 2016-06-30 LensingPaper_draft_v4.01.tex
Added one comment to referee report.
Finished referee report (almost) and removed all TO DOs from the Paper.

The JAM modelling code by \citet{Cap08} evaluates \OLD{Equation \eqref{eq:explicitLOSvelCap}}\NEW{this expression for $\langle v_\text{los}^2\rangle(x',y')$} numerically for a given luminous tracer and mass distribution MGE and a given inclination.
\Wilma{[TO DO: MAke sure M and N are introduced somewhere.]}
%--------------------------------------------

@@

\Wilma{[TO DO: Referee says: "In section 4.4.1, the NFW profile is fitted only to the inner data, and then is
shown to not provide a good model for the data beyond that point. This is not
entirely surprising, given that the outer data were not used to constrain the model!
What happens if the NFW profile is fitted to both small- and large- radius data at
once? (Perhaps it is impossible to find a good NFW model across the whole radius
range, but if this is the case, then it should be stated.)"]}

@@

%--------------------------------------------
Figure \ref{fig:modelB4_triangle} shows the posterior probability distribution (pdf) of the fit sampled with an MCMC. Overplotted are also the priors used to constrain the NFW halo. The mean parameters are summarized in Table \ref{tab:modelB4_bestfit}. We find that the best fit NFW halo strives to be more massive and with a higher concentration (due to a smaller $r_s$) than  proposed by the priors. The model also prefers  very negative velocity anisotropies. Both, the high halo concentration and low $\beta_z$, are needed to reproduce the central dip of the $v_\text{rms}$ curve. Figure \ref{fig:modelB4_models} illustrates the range of best fit models according to the extent of the pdf. The models fit the $v_\text{rms}$ data in the inner regions quite well (Figure \ref{fig:modelB4_vrms}) and are also consistent with the Einstein mass (see Figure \ref{fig:modelB4_enclMass}). The extrapolation of the model to larger radii however overestimates the data, does not exhibit a drop around $\sim 6''$ at all and seems to be therefore overall too massive.

We also fitted a model with a cored logarithmic DM halo. The cored halo models are in general slightly less massive than the NFW halo and therefore fit the outer regions of J1331 better. However, the density profile of the cored halo as well as the I-band light distribution within the plane drop as $\rho(r) \propto r^{-2}$. There is therefore a strong degeneracy between the stellar mass and the DM. Overall, we were not able to obtain tight constraints on the cored logarithmic halo.
%--------------------------------------------

@@

\subsubsection{\NEW{Discussion}}

\NEW{A fit to all nine available $v_\text{rms}$ data points would force the DM halo to be less massive. But with decreasing DM contribution in the inner regions the model approaches the mass-follows-light model in the center, that we already ruled out in Section \ref{sec:results_JAM_SB_MfL}. Allowing for free $\Upsilon_{*}(R')$, $\beta_z(R')$ and/or non-NFW $\rho_\text{halo}(r)$ profiles would over-fit the data again. Without further priors on the stellar $\Upsilon_{*}(R')$ and/or more data points, we can make only the following statement: A model with constant $\Upsilon_*$, moderately tangential $\beta_z$ and a spherical and massive DM component with high concentration can explain the central kinematics of J1331 very well, we have however ruled out that in this case the DM halo will follow a NFW halo profile at both small and large radii.}
%--------------------------------------------

@@

From kinematics along the photometric major axis only it \OLD{cannot}\NEW{can however not} be determined if such a twist or misalignment is present in J1331\NEW{, but it should be immediately visible in a 2D kinematic map}.\OLD{ But if it were, our modelling assumption of axisymmetry would not be satisfied anymore.} \OLD{We note that kinematics measured along an axis different from the kinematic major axis would result in $v_\text{rot}$ measurements being lower than the actual maximum $v_\text{circ}$. At large radii this could lead to an underestimation of the DM halo mass.}

\Wilma{[TO DO: Mention that warps would require other modelling techniques than the axisymmetric Jeans models.]}
%--------------------------------------------

@@

We also compare our results from Section \ref{sec:results_JAM_NFW} with the study by \citet{SWELLSV}. They found that the bulge of J1331 has an IMF \emph{more} bottom-heavy than the Salpeter IMF from fitting a NFW halo to (i) the Einstein mass and (ii) gas kinematics at larger radii $\gtrsim 8''$. In Section \ref{sec:results_JAM_NFW} we fitted a mass model with NFW halo to (i) the Einstein mass and (ii) stellar kinematics within $\sim 3.5''$. Our best fit $\Upsilon_\text{I,*} = 4.2 \pm 0.2$ (see Table \ref{tab:modelB4_bestfit}) indicates a \emph{less} bottom-heavy IMF than the Salpeter IMF. \NEW{In the following section we will argue why we do not think that the \cite{SWELLSV} model is a good model for the central regions of J1331. We therefore also do not believe their findings that the central IMF is more bottom-heavy than Salpeter.} \Wilma{[TO DO: Find new place for: "\cite{SWELLSV} found systematically lower NFW halo masses ($v_\text{circ,halo}(5'') \sim 120~\text{km s}^{-1}$ according to their Figure 2) than we did ($v_\text{circ,halo}(R'=5'') \sim 200~\text{km s}^{-1}$)."]}
%--------------------------------------------

@@

%--------------------------------------------
\subsection{On J1331's \OLD{central }kinematics}
%--------------------------------------------

%--------------------------------------------
\NEW{Overall, the kinematics of the merger remnant J1331 are peculiar. In particular there are two features in the $v_\text{rms}$ curve that are hard to explain with standard modelling techniques. The first feature, the deep central $v_\text{rms}$ dip, was studied in detail in this work. The second feature, the drop and rise in $v_\text{rms}$ around \Wilma{[TO DO]}, is even harder to explain; below we will speculate about possible reasons that could cause such a signature.}

@@

\NEW{First, we will compare our modelling results with the modelling results by \citet{SWELLSV} within \Wilma{[TO DO]}.}

@@

\NEW{Around $R'\sim7''$ the rotation curve and dispersion show some wiggles which lead to a spontaneous rise of the $v_\text{rms}$. As the spiral arms cross the major axis around this radius, we suspect they cause this disturbance of the axisymmetric kinematics. The DM halo should also start to dominate the kinematics at larger radii, whose total mass we expect more of the order of what \citet{SWELLSV} found and which we suspect has a profile that deviates strongly from a NFW halo at both small and large radii. A more massive halo of NFW shape---as the one we found in this work---could however exist in J1331 if the galaxy had a strong kinematic warp in its disk due to the merger.}

>>>>

The JAM modelling code by \citet{Cap08} evaluates \OLD{Equation \eqref{eq:explicitLOSvelCap}}\NEW{this expression for $\langle v_\text{los}^2\rangle(x',y')$} numerically for a given luminous tracer and mass distribution MGE and a given inclination.
%--------------------------------------------

@@

@@

%--------------------------------------------
Figure \ref{fig:modelB4_triangle} shows the posterior probability distribution (pdf) of the fit sampled with an MCMC. Overplotted are also the priors used to constrain the NFW halo. The mean parameters are summarized in Table \ref{tab:modelB4_bestfit}. We find that the best fit NFW halo strives to be more massive and with a higher concentration (due to a smaller $r_s$) than  proposed by the priors. The model also prefers  very negative velocity anisotropies. Both, the high halo concentration and low $\beta_z$, are needed to reproduce the central dip of the $v_\text{rms}$ curve. Figure \ref{fig:modelB4_models} illustrates the range of best fit models according to the extent of the pdf. The models fit the $v_\text{rms}$ data in the inner regions quite well (Figure \ref{fig:modelB4_vrms}) and are also consistent with the Einstein mass (see Figure \ref{fig:modelB4_enclMass}). The extrapolation of the model to larger radii however overestimates the data, does not exhibit a drop around $\sim 6''$ at all and seems to be therefore overall too massive.
%--------------------------------------------

@@

\subsubsection{\NEW{Further tests and discussion}}

We also fitted a model with a cored logarithmic DM halo. The cored halo models are in general slightly less massive than the NFW halo and therefore fit the outer regions of J1331 better. However, the density profile of the cored halo as well as the I-band light distribution within the plane drop as $\rho(r) \propto r^{-2}$. There is therefore a strong degeneracy between the stellar mass and the DM. Overall, we were not able to obtain tight constraints on the cored logarithmic halo. \NOTE{[NOTE: this paragraph was moved from 4.4.2 to 4.4.4.]}

\NEW{As we only fitted the halo models to the inner regions, it is not surprising that they do not fit at larger radii. A fit to all nine available $v_\text{rms}$ data points would force the DM halo to be less massive. But with decreasing DM contribution in the inner regions the model approaches the mass-follows-light model in the center, that we already ruled out in Section \ref{sec:results_JAM_SB_MfL}. Allowing for free $\Upsilon_{*}(R')$, $\beta_z(R')$ and/or more flexible $\rho_\text{halo}(r)$ profiles, which would be the reasonable next step, would over-fit the data again. Without further priors on the stellar $\Upsilon_{*}(R')$ and/or more data points, we can make only the following statement: A model with constant $\Upsilon_*$, moderately tangential $\beta_z$ and a spherical and massive DM component with high concentration can explain the central kinematics of J1331 very well. We have however ruled out that in this case the DM halo will follow a NFW halo profile at both small and large radii.}

\NEW{We will also discuss the kinematics at larger radii in more depth in the discussion section \ref{sec:discussion_kinematics}.}

@@

From kinematics along the photometric major axis only it \OLD{cannot}\NEW{can however not} be determined if such a twist or misalignment is present in J1331\NEW{, but it should be immediately visible in a 2D kinematic map}.\OLD{ But if it were, our modelling assumption of axisymmetry would not be satisfied anymore.} \OLD{We note that kinematics measured along an axis different from the kinematic major axis would result in $v_\text{rot}$ measurements being lower than the actual maximum $v_\text{circ}$. At large radii this could lead to an underestimation of the DM halo mass.}
%--------------------------------------------

@@

We also compare our results from Section \ref{sec:results_JAM_NFW} with the study by \citet{SWELLSV}. They found that the bulge of J1331 has an IMF \emph{more} bottom-heavy than the Salpeter IMF from fitting a NFW halo to (i) the Einstein mass and (ii) gas kinematics at larger radii $\gtrsim 8''$. In Section \ref{sec:results_JAM_NFW} we fitted a mass model with NFW halo to (i) the Einstein mass and (ii) stellar kinematics within $\sim 3.5''$. Our best fit $\Upsilon_\text{I,*} = 4.2 \pm 0.2$ (see Table \ref{tab:modelB4_bestfit}) indicates a \emph{less} bottom-heavy IMF than the Salpeter IMF. \NEW{In the following section we will argue why we do not think that the \cite{SWELLSV} model is a good model for the central regions of J1331. We therefore also do not believe their findings that the central IMF is more bottom-heavy than Salpeter.}
%--------------------------------------------

@@

%--------------------------------------------
\subsection{On J1331's \OLD{central }kinematics} \label{sec:discussion_kinematics}
%--------------------------------------------

%--------------------------------------------
\NEW{Overall, the kinematics of the merger remnant J1331 are peculiar. In particular there are two features in the $v_\text{rms}$ curve that are hard to explain with standard modelling techniques. The first feature, the deep central $v_\text{rms}$ dip, was studied in detail in this work. The second feature, the drop and rise in $v_\text{rms}$ around $R'\sim6''$, is even harder to explain; below we will speculate about possible reasons that could cause such a signature.}

@@

\NEW{First, we will compare our modelling results with the modelling results by \citet{SWELLSV} within $R_\text{eff}$.}

@@

\NEW{Around $R'\sim7''$ the rotation curve and dispersion show some wiggles which lead to a spontaneous rise of the $v_\text{rms}$. As the spiral arms cross the major axis around this radius, we suspect they cause this disturbance of the axisymmetric kinematics.}

\NEW{\emph{(iii) The profile of the DM halo at $R'\gtrsim 5''$:}}

\NEW{The DM halo should start to dominate the kinematics at larger radii. \cite{SWELLSV}, who fitted a NFW halo to the gas kinematics in the outer regions, found lower halo masses ($v_\text{circ,halo}(5'') \sim 120~\text{km s}^{-1}$ according to their Figure 2) than we did ($v_\text{circ,halo}(5'') \sim 200~\text{km s}^{-1}$, Figure \ref{fig:vcirc_comparison}). As we did not fit the outer regions and only used a prior for $v_\text{200}$, their result in this regime is more reliable. Given our findings that a NFW halo does not fit the kinematics at both small and large radii, we suspect that the true halo has a profile that deviates strongly from a NFW halo. A more massive halo of NFW shape---as the one we found in this work---could however exist in J1331 if the galaxy had a strong kinematic warp in its disk due to the merger.}
